### 2015-04-21 tex/presentation/presentation.tex

\documentclass[usenames,dvipsnames]{beamer}
\usetheme{Boadilla}
\usepackage{subfig} 
\usepackage{multicol}
\usepackage{color}
\usepackage[english]{babel}

\title[Clustering Approx Algs]{Local Search Approximation Algorithms for Clustering}
\author{Dave Lingenbrink, Alice Paul, Calvin Wylie}
\date{April 23, 2015}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command: 
%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Motivation}
Look for structure in unlabeled data.
\begin{itemize}
\item Social networks
\item Handwritten letters
\item Grouping customer classes
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{$k$-Means Algorithm and Local Search}
Objective: partition the data into sets $S_1, S_2, \ldots, S_k$ with means $\mu_1, \ldots, \mu_k$ to minimize $\sum_{i=1}^k \sum_{x \in S_i} ||x-\mu_i||^2$.

\begin{center}
\includegraphics<1>[width=3.75in]{k-means-1}
\includegraphics<2>[width=3.75in]{k-means-2}
\hspace{0.6in}\includegraphics<3>[width=3in]{k-means-3}
\includegraphics<4>[width=3in]{k-means-4}
\end{center}
\end{frame}

\begin{frame}
\frametitle{$k$-Median Objective Function}
Objective: find a subset of centers $S \subseteq X$ such that $|S| \leq k$ to minimize $\sum_{x \in X} \min_{s \in S} d(x - s)$.

\begin{center}
\includegraphics<1>[width=3in]{k-median-1}
\includegraphics<2>[width=3in]{k-median-2}
\includegraphics<3>[width=3in]{k-median-3}
\includegraphics<4>[width=3in]{k-median-4}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Proposed Algorithms}
\begin{enumerate}
\item Multiswap Algorithm: look at local swaps of size $p$. 
\[ \left(3+ \frac{2}{p} \right) \text{-approximation ratio}\]
\item Minimizing a monotone decreasing supermodular function: look at local swaps using a potential function $\psi$.
\[ \left(1 + \frac{c}{1-c} \cdot e^{-1} + \frac{1}{1-c} \cdot O(\varepsilon) \right) \text{-approximation ratio (whp)}\]
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Hiearchical Clustering}
Look for layered structure in unlabeled data and finding a natural level of clustering.
\begin{center}
\includegraphics[width=2.7in]{hierarchical-clustering-example}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Agglomerative Clustering}
Repeatedly merge the two closest clusters according to some distance function $d(C_1, C_2)$. 
\begin{center}
\includegraphics[width=2in]{agglomerative-clustering}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Hiearchical $k$-Median Objective Function}
Find subsets $S_1, S_2, \ldots, S_n$ ($|S_i| = i$) and assignments $a_1, \ldots, a_n$ such that the clusters are nested:
\begin{enumerate}
\item $S_i \subseteq S_{i+1}$ for $i = 1, \ldots, n-1$,
\item $\forall j \in X$, if $a_{i+1}(j) \in S_i$, then $a_i(j) = a_{i+1}(j)$, and
\item $\forall j, k \in X$, if $a_{i+1}(j) = a_{i+1}(k)$, then $a_{i}(j) = a_{i}(k)$. 
\end{enumerate}

\vspace{0.5in}
There exists an approximation algorithm that given $k$-median solutions for $k=1, \ldots, n$, will generate a hierarchical clustering such that the $k$-median objective function on any level increases by at most a factor of 20.71 (or a factor of 10.03 in expectation when randomized).
\end{frame}

\begin{frame}
\frametitle{Proposed Algorithm}
\begin{center}
\includegraphics<1>[width=3in]{hierarchical-1}
\includegraphics<2>[width=3in]{hierarchical-2}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Data Sets and Preliminary Results}
\begin{enumerate}
\item Iris and soybean data sets from UCI Machine Learning Repository
\item Generated Gaussian data
\item OR Library's $p$-median instances
\end{enumerate}

Comparisons in adjusted mutual information (weighted for hierarchical results), computation time, and comparison to optimal $k$-median solutions.
\end{frame}

\begin{frame}
Questions?

\vspace{0.5in}
Learn more about the proposed algorithms:
\begin{itemize}
\item Arya et al. Local search heuristics for k-median and facility location problems. \emph{SIAM Journal on Computing}, 33(3):544-562, 2004.
\item Lin et al. A general approach for incremental approximation and hierarchical clustering. In \emph{Proceedings of the Seventeenth Annual ACM-SIAM Symposium on Discrete Algorithms}, SODA `06, pp.1147-1156, Philadelphia, PA, USA, 2006. 
\item Sviridenko, Maxim and Ward, Justin. Tight bounds for submodular and supermodular optimization with bounded curvature. arXiv:1311.4728v3 [cs.DS] 2014.
\end{itemize}
\end{frame}

\end{document}